---
title: "Bigfoot: Fact or Non-Fiction"
---


```{r, message=FALSE, warning=FALSE, include=FALSE}
library(leaflet)
library(dplyr)
library(plotly)
```

```{r, echo=FALSE, warning=FALSE}
dat <- read.csv("C:/Users/ATrac/OneDrive/Documents/Web/atraczyk.github.io/bfro_reports_geocoded.csv")


# geo styling
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showland = TRUE,
  landcolor = "grey",
  subunitcolor = "grey",
  countrycolor = "grey",
  countrywidth = 0.5,
  subunitwidth = 0.5
)

p <- plot_geo(dat, lat = ~latitude, lon = ~longitude) %>%
  add_markers(
    text = ~paste(title, classification, sep = "<br />"),
    symbol = I("circle-dot"), size = I(3), hoverinfo = "text"
  ) %>%
  layout(
    title = 'Bigfoot Sightings United States', geo = g
  )

p
```

As populations continue to expand into previously rural areas, we find ourselves face to face with what lives in the deep of the woods. In this project, for the first time, we apply machine learning and super-duper advanced visualization techniques to finally answer the question: Bigfoot, fact or non-fiction.

The study includes various analysis, briefly described below (click links, or use navbar on top of page to read futher):

### [Sub-Species Analysis](https://atrac2k.github.io/Portfolio/sub.html)

An analysis of regional patterns of Bigfoot through unsupervised learning, and general characteristics of these sub-species

### [Frequency Analysis](https://atrac2k.github.io/Portfolio/ts.html)

Analysis on the frequency of Bigfoot sightings, including predicting future sighting frequency

### [About the Author](https://atrac2k.github.io/Portfolio/about.html)

Learn more about the cool guy who wrote this (and is actually like 80% sane)