---
title: "Sub-Species Analysis"
output: html_document
---

As most animals with a range the size of the United States tend to have more specific species within smaller areas (for instance, black, brown, or grizzly bears), it's likely to be more beneficial to specify smaller Bigfoot ranges. Some theories already propose such sub-species of Bigfoot exist (Sasquatch in the Pacific Northwest, Grassman in Ohio, Skunk Ape in Florida). By finding these ranges, we should expect each sub-species to act somewhat differently.

To extract this information from our data of Bigfoot sightings, we turn to unsupervised learning, and specifically Expectation Maximization (EM). We use expectation maximization over other clustering techniques (such as K-Mean's, or DBSCAN) for three reasons: 

  * EM clustering is referred to as "soft-margin" clustering. In EM this means each point of data is calculated as a probability of lying in each cluster. We can therefore tweak the model to remove outliers, and center in on Bigfoot ranges.
  * EM works much better than other techniques on data with different densities. As seen in the map at the top of the page, the density of Bigfoot sightings is quite different from one region to another.
  * Other clustering techniques tend to assume each cluster should have an equal number of data points within them. This can create clusters which are awkwardly shaped to make the clusters even
  
Now that we have a technique, we need to choose just how many classes are in our EM model (classes are the equivalent of sub-species). To choose the optimum number of classes, we perform cross-validation on class sizes from 1 to 10. We choose to evaluate the performance of these models by Bayesian Information Criterion (BIC). We choose this metric over Akaike Information Criterion (AIC) as it has a larger penalty for larger models, and will help prevent us from over fitting the model. The code and results are below:

```{r, message=FALSE, warning=FALSE, include=TRUE,eval=FALSE}
library(EMCluster)
library(ggplot2)

##Read Data
dat <- read.csv("C:/Users/ATrac/OneDrive/Documents/Web/atraczyk.github.io/bfro_reports_geocoded.csv")
dat <- dat[!is.na(dat$latitude)&!is.na(dat$longitude),]

##Setup for loop
set.seed(0)
temp <-  seq(1:10)
cvresults <- data.frame(temp)

##for loop to run various sizes of cross validation
for (j in 1:2){
  for (i in 1:10){
    test <- simple.init(dat[,c("latitude","longitude")],nclass=i)
    test <- shortemcluster(dat[,c("latitude","longitude")],test)
    test2 <- emcluster(dat[,c("latitude","longitude")],test,assign.class = T)
    temp[i] <- summary(test2)$BIC
  }
  cvresults <- data.frame(cvresults,temp)
}

##Calculate mean,min,max of each class
cvresults <- cvresults[,-1]
cvresults[,"average"] <- rowMeans(cvresults,na.rm=T)
for (i in 1:nrow(cvresults)){
  cvresults[i,"min"] <- min(cvresults[i,],na.rm=T)
  cvresults[i,"max"] <- max(cvresults[i,],na.rm=T)
}
cvresults[,"classes"] <- 1:10
```
```{r, message=FALSE, warning=FALSE, include=FALSE}
library(EMCluster)
library(ggplot2)
library(png)
library(grid)
cvresults <- read.csv("C:/Users/ATrac/OneDrive/Documents/cvresults.csv")

bigfootpng <- download.file("http://images.gawker.com/195g0l3khm7z0png/c_fit,fl_progressive,q_80,w_636.png",destfile = 'bigfoot.png',mode='wb')
bigfootpng <- readPNG("bigfoot.png")
g <- rasterGrob(bigfootpng, interpolate=TRUE) 
```
```{r}
##Plot cross validation results
ggplot(data=cvresults, aes(x=classes,y=average))+
  annotation_custom(rasterGrob(bigfootpng,width = unit(1,"npc"),  height = unit(1,"npc")),  -Inf, Inf, -Inf, Inf) +
  geom_errorbar(aes(ymin=min,ymax=max,col='red'),show.legend=F,width=0)+
  geom_line()+
  geom_point() + 
  geom_line(aes(x=classes,y=min,col='red'),show.legend=F)+
  geom_line(aes(x=classes,y=max,col='red'),show.legend=F)+
  labs(y="BIC",title="Cross Validation Results, with error")+
  scale_x_continuous(name="Classes", breaks=c(1:10))
```

From our cross validation, I chose 7 Classes. BIC was small, and larger classes started to see growing error bars (implying over fitting). We now calculate our EM model with 7 classes, and plot the results below:
```{r, message=FALSE, warning=FALSE, include=FALSE}
dat <- read.csv("C:/Users/ATrac/OneDrive/Documents/Web/atraczyk.github.io/bfro_reports_geocoded.csv")
dat <- dat[!is.na(dat$latitude)&!is.na(dat$longitude),]
```
```{r, message=FALSE, warning=FALSE}
set.seed(0)
test <- simple.init(dat[,c("latitude","longitude")],nclass=7)
test <- shortemcluster(dat[,c("latitude","longitude")],test)
test2 <- emcluster(dat[,c("latitude","longitude")],test,assign.class = T)

emclass <- dat[,c("latitude","longitude")]
emclass[,"Class"] <- as.factor(test2$class)
```
```{r, message=FALSE, warning=FALSE, include=FALSE}
emclass[,"Species"] <- ""
for (i in 1:nrow(emclass)){
  if (emclass[i,3]==1){
    emclass[i,4] <- "Mogollon"
  }
    if (emclass[i,3]==2){
    emclass[i,4] <- "Grassman"
    }
    if (emclass[i,3]==3){
    emclass[i,4] <- "Momo"
    }
    if (emclass[i,3]==4){
    emclass[i,4] <- "Wendigo"
    }
    if (emclass[i,3]==5){
    emclass[i,4] <- "Skunk Ape"
    }
    if (emclass[i,3]==6){
    emclass[i,4] <- "Sasquatch"
    }
    if (emclass[i,3]==7){
    emclass[i,4] <- "Grandis Alekus"
  }
  
}
```
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(plotly)

g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showland = TRUE,
  landcolor = "grey",
  subunitcolor = "grey",
  countrycolor = "grey",
  countrywidth = 0.5,
  subunitwidth = 0.5
)

p <- plot_geo(emclass, lat = ~latitude, lon = ~longitude) %>%
  add_markers(
    color = ~Species, symbol = I("square"), size = I(4)
  ) %>%
  layout(
    title = 'Bigfoot Sub-Species', geo = g
  )

p
```

These seven sub-species of Bigfeet seem to follow regional differences already used in most descriptions:

  * Along the West Coast, Sasquatch
  * In Arizona, the Mogollon Monster
  * In the Midwest, Missouri Monster (Momo)
  * Around Lake Superior, Wendigo
  * Grassman in Ohio
  * Skunk Ape in Florida
  
Our seventh sub-species of Bigfoot seem to have no regional equivalent (I got bored after checking like 2 Google results). Thus, our EM model must have found an entirely new sub-species of Bigfoot in the Southeastern states. As the discoverer of this species, I have chosen to name is "Grandis Alekus".

### Sub-Species Characteristics

Now that we have our various sub-species, we can start to study each individually.
```{r, message=FALSE, warning=FALSE, include=FALSE}
library(lubridate)
dat <- read.csv("C:/Users/ATrac/OneDrive/Documents/Web/atraczyk.github.io/bfro_reports_geocoded.csv")
dat <- dat[!is.na(dat$latitude)&!is.na(dat$longitude),]

dat[,"Class"] <- as.factor(test2$class)
dat[,"Species"] <- ""
for (i in 1:nrow(dat)){
  if (dat[i,28]==1){
    dat[i,29] <- "Mogollon"
  }
    if (dat[i,28]==2){
    dat[i,29] <- "Grassman"
    }
    if (dat[i,28]==3){
    dat[i,29] <- "Momo"
    }
    if (dat[i,28]==4){
    dat[i,29] <- "Wendigo"
    }
    if (dat[i,28]==5){
    dat[i,29] <- "Skunk Ape"
    }
    if (dat[i,28]==6){
    dat[i,29] <- "Sasquatch"
    }
    if (dat[i,28]==7){
    dat[i,29] <- "Grandis Alekus"
  }
  
}
dat[,"Month"] <- month(as.Date(dat$date,"%Y-%m-%d"))
```
```{r}
qplot(Month,data=dat,facets=Species~.,geom="histogram",fill=Species,binwidth=1,main="Sightings By Month")+
  scale_x_continuous(name="Month", breaks=c(1:12),labels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
```

#### Climate

We take note from the above plot of sightings by month, that in most sub-species of Bigfoot sightings drop during the winter months (notable exceptions include Grandis Alekus, Grassman, Mogollon). It is my hypothesis that most Bigfoot sub-species hibernate during winter, and thus sightings drop off (of course it isn't because less people are outside during the winter. Don't be preposterous). We test this hypothesis by instead plotting by the low temperature during each sighting:

```{r, message=FALSE, warning=FALSE}
qplot(round(temperature_low,0),data=dat,facets=Species~.,geom="histogram",fill=Species,binwidth=2,main="Sightings By Low Temperature")+
  scale_x_continuous(name="Low Temperature")
```

When viewing the data by low temperature, we notice that sightings when the temperature is blow freezing (32) tend to become far less frequent. This does give quite the credibility to our hibernation theory. This seems to also closely resemble the hibernation patterns of bears in North America, which tend to begin hibernation just as temperatures are nearing freezing.

#### Diet

Many studies also assume Bigfoot has a similar diet to bears. Although little is known about the diet of Bigfoot, it shouldn't be much of a stretch to assume Grandis Alekus' favorite food is cut up hot dogs and ketchup (just like their namesake). Since I'm too lazy to find enough data to test this hypothesis, I'm assuming I'm correct because I haven't seen data to prove otherwise. 

#### Migration Patterns

Since we have geocoded data for reported sightings, we can also test the average position of sightings by month. As we proved that Bigfoot hibernates over winter, we can expect Bigfoot to not show any major migratory patterns.

```{r}
species <- c("Mogollon","Grassman","Momo","Wendigo","Skunk Ape","Sasquatch","Grandis Alekus")
dat[,"Month"] <- as.numeric(substr(dat$date,6,7))

migration <- data.frame(NULL)
migration[1:12,1] <- species[1]
migration[13:24,1] <- species[2]
migration[25:36,1] <- species[3]
migration[37:48,1] <- species[4]
migration[49:60,1] <- species[5]
migration[61:72,1] <- species[6]
migration[73:84,1] <- species[7]

migration[,2] <- seq(1,12)
migration[,3:4] <- NA

for (i in 1:nrow(migration)){
  temp <- filter(dat,Species == migration[i,1] & Month == migration[i,2])
  migration[i,3] <- mean(temp$latitude)
  migration[i,4] <- mean(temp$longitude)
}

colnames(migration) <- c("Species","Month","Lat","Long")
```
```{r}
library(leaflet)

leaflet() %>%
  addTiles() %>%
  addPolylines(lat = migration[migration[,1]=="Mogollon",3],lng=migration[migration[,1]=="Mogollon",4],color="blue",opacity=1) %>%
  addPolylines(lat = migration[migration[,1]=="Grassman",3],lng=migration[migration[,1]=="Grassman",4],color="green",opacity=1) %>%
  addPolylines(lat = migration[migration[,1]=="Momo",3],lng=migration[migration[,1]=="Momo",4],color="coral",opacity=1) %>%
  addPolylines(lat = migration[migration[,1]=="Wendigo",3],lng=migration[migration[,1]=="Wendigo",4],color="yellow",opacity=1) %>%
  addPolylines(lat = migration[migration[,1]=="Skunk Ape",3],lng=migration[migration[,1]=="Skunk Ape",4],color="purple",opacity=1) %>%
  addPolylines(lat = migration[migration[,1]=="Sasquatch",3],lng=migration[migration[,1]=="Sasquatch",4],color="brown",opacity=1) %>%
  addPolylines(lat = migration[migration[,1]=="Grandis Alekus",3],lng=migration[migration[,1]=="Grandis Alekus",4],color="gold",opacity=1) %>%
  addLegend("bottomleft",colors=c("blue","green","coral","yellow","purple","brown","gold"),labels=species,opacity=1,title="Bigfoot Migration Patterns")

```

We notice that none of the seven Bigfoot species exhibit any major migratory patterns. What we notice instead, is that movement patterns of Bigfoot seem to be highly restricted by populated areas. More "remote" Bigfoot ranges (such as those of Sasquatch, Mogollon, or Wendigo) tend to have greater movement patterns than their cousins in more populous areas. The most restricted range seems to be that of the Skunk Ape, which likely can't move freely without being accosted by some retiree. 

Again, relating back to bears, we have actually seen similar patterns in their movements. Studies found that bear ranges were greatly restricted by the presence of highways, which led to the creation of the Yellowstone to Yukon Conservation Initiative; creating free movement corridors for bears through the use of  wildlife crossings. The author proposes a similar program, the "No Bigfoot Left Behind Initiative" to remove the movement restrictions we humans have created for Bigfoot.